---
title: "Accident Severity in USA"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Makes sure that the required packages are properly installed; if they are not installed, R helps to install them
# Retrieved on 04/19/25 from - https://stackoverflow.com/questions/4090169/elegant-way-to-check-for-missing-packages-and-install-them
required_packages <- c("tidyverse", "glmnet", "patchwork", "caret", "scales", "gifski", "gganimate", "tidymodels", "MASS", "car",
                       "pROC", "doParallel")
new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) {
  install.packages(new_packages)
}


library(tidyverse)
library(patchwork)
library(glmnet)
library(caret)
library(scales)
library(gganimate)
library(gifski)
library(tidymodels)
library(MASS)
library(car)
library(pROC)
library(doParallel)


my_theme <- theme_minimal() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, color = "black", face = "plain"),
    title = element_text(size = 14, color = "black", face = "bold"),
    legend.text = element_text(size = 12, color = "black"),
    legend.title = element_text(
      color = "black",
      size = 12, margin =
        ggplot2::margin(t = 0, r = 20, b = 0, l = 0, unit = "pt")
    ),
    strip.text = element_text(size = 14, color = "black"),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "bottom",
    panel.spacing = unit(15, "pt"),
    axis.line = element_line(colour = "black"),
    panel.grid = element_line(color = "#f3f6f4")
  )

theme_set(my_theme)

options(warn = -1)
```


```{r}
# Read data - Accidents from 2018 -> 2022

accidents <- read_csv("US_Accidents.csv")

# Check NA values
sapply(accidents, function(y) sum(length(which(is.na(y)))))

```


## Part 1: Understanding the Dataset


#### A: Temporal Trend

```{r, fig.height = 6, fig.width = 10}
quarter_arrangement <- c("Q1 2018", "Q2 2018", "Q3 2018", "Q4 2018",
                         "Q1 2019", "Q2 2019", "Q3 2019", "Q4 2019",
                         "Q1 2020", "Q2 2020", "Q3 2020", "Q4 2020",
                         "Q1 2021", "Q2 2021", "Q3 2021", "Q4 2021",
                         "Q1 2022", "Q2 2022", "Q3 2022", "Q4 2022")

time_trend1 <- accidents %>% 
  count(quarter_date) %>% 
  mutate(quarter_date = factor(quarter_date, levels = quarter_arrangement)) %>% 
  arrange(quarter_date) %>% 
  mutate(is_q4 = grepl("Q4", quarter_date))

time_trend1 %>% 
  ggplot(aes(x = quarter_date, y = n)) +
  geom_line(group = 1, color = "#C5050C") +
  geom_point() +
  geom_point(data = filter(time_trend1, is_q4),
             aes(x = quarter_date, y = n, color = "Quarter 4"),
             shape = 21, size = 5, stroke = 1.5) +
  labs(x = "", y = "Number of Accidents", title = "Quarterly Trend of Accidents (Fatal/Non-fatal), 2018 - 2022",
       color = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_blank(),
        axis.ticks = element_line(color = "black", size = 0.5),
        legend.position = c(0.1, 0.9)) +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  scale_x_discrete(breaks = quarter_arrangement[seq(1, length(quarter_arrangement), by = 2)]) +
  scale_color_manual(values = c("Quarter 4" = "darkblue")) +
  annotate("segment", x = 12.25, y = 180000, xend = 11.25, yend = 160000,
         arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("segment", x = 10.6, y = 460000, xend = 11.6, yend = 480000,
         arrow = arrow(type = "closed", length = unit(0.02, "npc"))) +
  annotate("text", x = 15, y = 190000, label = "Increased caution/tightened restrictions\ndo to rising COVID numbers") +
  annotate("text", x = 10, y = 435000, label = "Election + holiday +\npandemic fatigue")

ggsave("quarterly_trend.svg")
  
```


#### B: Geographic Trend

```{r, fig.height = 6, fig.width = 10}
us_states <- map_data("state")

census_pop <- read_csv("Census_Population.csv") %>%
  pivot_longer(
    cols        = starts_with("pop_"),
    names_to    = "year",
    names_prefix= "pop_",
    values_to   = "population"
  ) %>%
  mutate(year = as.integer(year))

geog_accidents <- accidents %>% 
  count(year_date, State) %>% 
  left_join(census_pop, by = c("State" = "abbr", "year_date" = "year")) %>% 
  mutate(accidents_per_100k = round((n / population) * 100000)) %>% 
  mutate(bins = case_when(
    accidents_per_100k <= 100 ~ "100 or less",
    accidents_per_100k <= 200 ~ "101 - 200",
    accidents_per_100k <= 300 ~ "201 - 300",
    accidents_per_100k > 300 ~ "301 or more"
  )) %>%
  mutate(bins = forcats::fct(bins, levels = c(
    "100 or less", "101 - 200",
    "201 - 300", "301 or more"
  ))) %>% 
  mutate(state = tolower(state))

base_map <- us_states %>%
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", fill = NA) +
  theme(
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  labs(x = "", y = "",
       title = "Accident Rate per Capita")


accidents_over_time <- us_states %>% 
  dplyr::select(-subregion) %>% 
  arrange(region) %>%  
  inner_join(geog_accidents, by = c("region" = "state"))

num_years <- max(accidents$year_date) - min(accidents$year_date) + 1


label_df <- accidents_over_time %>%
  group_by(State, year_date) %>%
  summarize(
    long = mean(range(long)),
    lat  = mean(range(lat)),
    .groups = "drop"
  ) %>% 
  mutate(lat =  ifelse(State == "ID", lat - 1, lat),
         long = ifelse(State == "MI", long + 1.5, long),
         lat =  ifelse(State == "MI", lat - 1.5, lat),
         long = ifelse(State == "MN", long - 1, long),
         lat =  ifelse(State == "MN", lat - 1, lat),
         long = ifelse(State == "LA", long - 1, long),
         long = ifelse(State == "FL", long + 2.5, long),
         lat =  ifelse(State == "KY", lat - 0.5, lat),
         long = ifelse(State == "NC", long + 0.5, long),
         lat =  ifelse(State == "NC", lat + 0.5, lat),
         long = ifelse(State == "VA", long + 1.5, long),
         lat =  ifelse(State == "VA", lat - 0.5, lat),
         long = ifelse(State == "MA", long - 0.5, long),
         lat =  ifelse(State == "MA", lat + 0.5, lat),
         long = ifelse(State == "VT", long - 0.25, long),
         lat =  ifelse(State == "MD", lat + 0.5, lat))

animation <- base_map +
  geom_polygon(data = accidents_over_time, aes(
    x = long, y = lat, group = group,
    fill = bins
  ), color = "black") +
  geom_text(
    data = label_df,
    aes(x = long, y = lat, label = State),
    size = 3.5,
    color = "black",
    inherit.aes = FALSE
  ) + 
  scale_fill_manual(values = rev(c("#EA4C46", "#F07470", "#F6BDC0", "#FFEBEA"))) +
  labs(fill = "") +
  transition_time(year_date) +
  ggtitle("Accidents (Fatal/Non-fatal) per 100,000 People by State",
    subtitle = "Year: {frame_time}"
  )
  

animate(animation,
        nframes = num_years, fps = 0.8, renderer = gifski_renderer("accidents_rate.gif"), height = 1080, width = 1920,
        res = 175
)

```

```{r, fig.height = 5, fig.width = 12}

p <- list()

city_accidents <- accidents %>% 
  drop_na(City) %>% 
  count(City) %>% 
  arrange(desc(n))

street_accidents <- accidents %>% 
  drop_na(Street) %>% 
  mutate(Street = str_remove(Street, " [NSWE]$")) %>% 
  count(Street) %>% 
  arrange(desc(n))


p[[1]] <- city_accidents %>% 
  head(10) %>% 
  ggplot(aes(y = reorder(City, n), x = n)) +
  geom_col(color = "black", fill = "orangered2", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0), labels = label_number(scale = 1e-3, suffix = "K")) +
  labs(y = "", x = "Number of Accidents", 
       title = "Cities") +
  theme(
    panel.grid.major.y = element_blank()
  )


p[[2]] <- street_accidents %>% 
  head(10) %>% 
  ggplot(aes(y = reorder(Street, n), x = n)) +
  geom_col(color = "black", fill = "deepskyblue4", alpha = 0.8) +
  scale_x_continuous(expand = c(0, 0), labels = label_number(scale = 1e-3, suffix = "K")) +
  labs(y = "", x = "Number of Accidents", 
       title = "Roads") +
  theme(
    panel.grid.major.y = element_blank()
  )
  

p[[1]] + p[[2]] +
  plot_annotation(title = "Most Accident-prone Cities and Roads, 2018 - 2022")


ggsave("accident_city_roads.svg")

```


## Part 2: Severity Analysis

```{r, fig.height = 5, fig.width = 9}

p2 <- list()

p2[[1]] <- accidents %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  count(Severity) %>%
  mutate(total = sum(n),
         proportion = round(n / total, 2)) %>% 
  ggplot() +
  geom_col(aes(x = Severity, y = n, fill = Severity), show.legend = F, color = "black") +
  labs(x = "Severity", y = "Number of Accidents", fill = "",
       title = "Distribution of Accidents by Severity, 2018 - 2022",
       subtitle = "Severity: 1 = Least Severe, 4 = Most Severe") +
  scale_fill_manual(values = rev(c("#EA4C46", "#F07470", "#F6BDC0", "#FFEBEA"))) +
  scale_y_continuous(labels = label_number(scale = 1e-6, suffix = "M"), expand = c(0, 0.01, 0.05, 0.1)) +
  theme(panel.grid.major.x = element_blank(),
        plot.subtitle = element_text(face = "plain")) +
  geom_text(aes(x = Severity, y = (n + 2e5), label = paste0(proportion * 100, "%")),
            size = 4)


p2[[1]]

ggsave("severity_distrib.svg")

```


```{r}
# Sample 50,000 observations from each Severity class
set.seed(2025)

accidents_trunc <- accidents %>% 
  drop_na() %>% 
  mutate(Severity = factor(Severity)) %>% 
  group_by(Severity) %>% 
  slice_sample(n = 50000,
               replace = F) %>% 
  ungroup() %>% 
  mutate(Severity = ifelse(Severity %in% c(1,2), "Minor", "Severe"),
         Severity = factor(Severity, levels = c("Minor", "Severe")))
```


#### A: Figures to Note: Year and Distance

```{r, fig.height = 5, fig.width = 9}
p2[[2]] <- accidents_trunc %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  group_by(year_date, Severity) %>% 
  summarize(counts = n()) %>% 
  group_by(year_date) %>% 
  mutate(total = sum(counts)) %>% 
  ungroup() %>% 
  mutate(proportion = round((counts/total) * 100, 2)) %>% 
  ggplot(aes(x = year_date, y = proportion, fill = Severity)) +
  geom_col(position = "stack", color = "black") +
  labs(x = "Year", y = "Proportion", 
       title = "Proportion of Accident Severity by Year") +
  scale_y_continuous(labels = label_number(suffix = "%"), expand = c(0, 0.01, 0.11, 0.1)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0"))) +
  theme(plot.subtitle = element_text(face = "plain"),
        panel.grid.major.x = element_blank())
  
p2[[2]]

ggsave("year_distrib.svg")

```


```{r, fig.height = 5, fig.width = 9}
accidents_trunc

accidents_trunc %>% 
  drop_na(`Distance(mi)`) %>% 
  ggplot(aes(x = Severity, y = log(`Distance(mi)`), fill = Severity)) +
  geom_boxplot(show.legend = F) +
  theme(panel.grid.major.x = element_blank()) +
  labs(y = "log(Distance (mi))",
       title = "Distribution of Impacted Road Length by Accident Severity") +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))

ggsave("distance_distrib.svg")

# Significant, but distance is a consequence of severity
t.test(accidents_trunc %>% filter(Severity == "Severe") %>%  pull(`Distance(mi)`),
       accidents_trunc %>% filter(Severity == "Minor") %>%  pull(`Distance(mi)`))


```


#### B: Relationship between Time (Period of the Day / Day of the Week / Month) and Accident Risk 

```{r, fig.height = 5, fig.width = 12}
# https://www.britannica.com/dictionary/eb/qa/parts-of-the-day-early-morning-late-morning-etc

time_accident <- accidents_trunc %>% 
  mutate(
    period = hour(Start_Time)) %>% 
  mutate(period = case_when(
    period >= 5 & period <= 11 ~ "Morning",
    period >= 12 & period <= 16 ~ "Afternoon",
    period >= 17 & period <= 20 ~ "Evening",
    period >= 21 | (period <= 4) ~ "Night"
  )) %>% 
  mutate(period = factor(period, 
                         levels = c("Morning", "Afternoon", "Evening", "Night"))) %>% 
  mutate(day_date = wday(Start_Time, abbr = T, label = T)) %>% 
  mutate(day_date = factor(day_date, levels = c("Mon", "Tue", "Wed", "Thu", "Fri",
                                                "Sat", "Sun"))) %>% 
  mutate(month_date = str_extract(month_date, "^[A-Za-z]+")) %>% 
  mutate(month_date = factor(month_date, 
                         levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                                    "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>% 
  dplyr::select(period, day_date, month_date, Severity)

period_graph <- time_accident %>% 
  group_by(period, Severity) %>% 
  summarize(counts = n()) %>% 
  group_by(period) %>% 
  mutate(total = sum(counts)) %>% 
  ungroup() %>% 
  mutate(proportion = round((counts/total), 2))


# Period
p2[[3]] <- ggplot(period_graph, aes(x = period, y = proportion, fill = Severity)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  labs(
    title = "By Period of the Day",
    x     = "",
    y     = "Proportion",
    fill  = "Severity\n"
  ) +
  scale_y_continuous(expand = c(0, 0, 0.1, 0), labels = scales::percent_format()) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0"))) +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14),
        legend.position = "right")

```


```{r, fig.height = 5, fig.width = 9}

day_graph <- time_accident %>% 
  mutate(day_date = case_when(
    day_date %in% c("Mon", "Tue", "Wed", "Thu", "Fri") ~ "Weekday",
    day_date %in% c("Sat", "Sun") ~ "Weekend"
  )) %>% 
  group_by(day_date, Severity) %>% 
  summarize(counts = n()) %>% 
  group_by(day_date) %>% 
  mutate(total = sum(counts)) %>% 
  ungroup() %>% 
  mutate(proportion = round((counts/total), 2)) %>% 
  mutate(proportion = ifelse(day_date == "Weekend" & Severity == 2, proportion + 0.01, proportion))

p2[[4]] <- day_graph %>% 
 ggplot(aes(x = day_date, y = proportion, fill = Severity)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  labs(x = "", y = "", 
       title = "By Parts of the Week",
       fill  = "Severity\n") +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0, 0.0, 0.01, 0.03)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0"))) +
  theme(plot.subtitle = element_text(face = "plain"),
        plot.title = element_text(size = 14),
        panel.grid.major.x = element_blank(),
        legend.position = "right")

```


```{r, fig.height = 5, fig.width = 12}

season_graph <- time_accident %>% 
  mutate(season = case_when(
    month_date %in% c("Dec", "Jan", "Feb") ~ "Winter",
    month_date %in% c("Mar", "Apr", "May") ~ "Spring",
    month_date %in% c("Jun", "Jul", "Aug") ~ "Summer",
    month_date %in% c("Sep", "Oct", "Nov") ~ "Fall"
  )) %>% 
  mutate(season = factor(season, levels = 
                           c("Spring", "Summer", "Fall",
                             "Winter"))) %>% 
  group_by(season, Severity) %>% 
  summarize(counts = n()) %>% 
  group_by(season) %>% 
  mutate(total = sum(counts)) %>% 
  ungroup() %>% 
  mutate(proportion = round((counts/total), 2)) %>% 
  mutate(proportion = ifelse(season == "Fall" & Severity == 2, proportion + 0.01, proportion))


p2[[5]] <- season_graph %>% 
 ggplot(aes(x = season, y = proportion, fill = Severity)) +
  geom_col(position = position_dodge(width = 0.9), color = "black") +
  labs(x = "", y = "", 
       title = "By Season",
       fill  = "Severity\n") +
  scale_y_continuous(labels = scales::percent_format(), expand = c(0, 0.0, 0.01, 0.03)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0"))) +
  theme(plot.subtitle = element_text(face = "plain"),
        plot.title = element_text(size = 14),
        panel.grid.major.x = element_blank(),
        legend.position = "right")

```


```{r, fig.height = 5, fig.width = 14}
(p2[[3]] + p2[[4]] + p2[[5]]) +
  plot_annotation(title = "Relationship between Time and Accident Risk") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("time.svg")
```


#### C: Effects of Weather Features on Accident Outcomes


```{r, fig.height = 5, fig.width = 9}

accidents_trunc <- accidents_trunc %>%
  mutate(Wind_Direction = case_when(
      Wind_Direction %in% c("N", "North", "NNE", "NNW", "NE", "NW") ~ "North",
      Wind_Direction %in% c("E", "East", "ENE", "ESE") ~ "East",
      Wind_Direction %in% c("S", "South", "SSW", "SW", "SE", "SSE") ~ "South",
      Wind_Direction %in% c("W", "West", "WSW", "WNW") ~ "West",
      Wind_Direction == "CALM" ~ "Calm",
      Wind_Direction %in% c("VAR", "Variable") ~ "Variable"
    )
  ) %>% 
  mutate(Wind_Direction = factor(Wind_Direction, levels = c("North", "South", "East", 
                                                            "West", "Calm", "Variable"))) %>% 
   mutate(
    Weather_Condition = case_when(
      Weather_Condition %in% c("Fair", "Clear", "Fair / Windy") ~ "Clear",
      Weather_Condition %in% c("Cloudy", "Mostly Cloudy", "Partly Cloudy", "Mostly Cloudy / Windy",
                               "Partly Cloudy / Windy", "Cloudy / Windy", "Scattered Clouds", "Overcast") ~ "Cloudy",
      Weather_Condition %in% c("Fog", "Haze", "Mist", "Smoke", "Fog / Windy", "Smoke / Windy", "Partial Fog",
                               "Patches of Fog / Windy", "Light Freezing Fog", "Widespread Dust", 
                               "Drizzle and Fog", "Patches of Fog", "Shallow Fog", "Blowing Dust / Windy", 
                               "Haze / Windy", "Sand / Dust Whirlwinds", "Blowing Dust") ~ "Fog/Haze/Dust",
      Weather_Condition %in% c("Light Rain", "Thunder", "Rain", "Thunder in the Vicinity", 
                               "Light Rain with Thunder", "Light Drizzle", "Drizzle",
                               "Rain / Windy", "Light Drizzle / Windy", "Light Freezing Rain",
                               "Light Rain / Windy", "N/A Precipitation", "Light Rain Shower",
                               "Light Freezing Drizzle", "Thunder / Windy", "Light Freezing Rain / Windy", "Freezing Drizzle",
                               "Light Thunderstorms and Rain") ~ "Rain",
      Weather_Condition %in% c("Heavy T-Storm", "T-Storm", "Heavy T-Storm / Windy", "Heavy Rain", "Showers in the Vicinity",
                               "T-Storm / Windy", "Heavy Rain / Windy", "Heavy Drizzle", "Freezing Rain",
                               "Rain Shower", "Heavy Rain Shower", "Heavy Drizzle", "Thunderstorms and Rain") ~ "Heavy Rain",
      Weather_Condition %in% c("Light Snow", "Snow", "Light Sleet", "Light Snow / Windy", "Snow / Windy", "Light Ice Pellets", 
                               "Light Sleet / Windy", "Light Snow Shower", 
                               "Light Blowing Snow", "Light Snow and Sleet", "Small Hail", "Light Thunderstorms and Snow") ~ "Snow",
      Weather_Condition %in% c("Heavy Snow / Windy", "Heavy Snow", "Blowing Snow", "Blowing Snow / Windy", "Sleet",
                               "Heavy Sleet", "Snow and Sleet", "Snow and Sleet / Windy", "Ice Pellets", 
                               "Heavy Snow with Thunder", "Squalls / Windy", "Hail") ~ "Heavy Snow",
      TRUE ~ "Other"))

```


##### 1: Wind Direction and Speed

```{r, fig.height = 5, fig.width = 12}

library(openair)

wind_rose <- accidents_trunc %>% 
  dplyr::select(Wind_Direction, `Wind_Speed(mph)`, Severity) %>%
  mutate(
    ws = `Wind_Speed(mph)`,
    # 2) map each cardinal label (and “Calm”/“Variable”) to degrees or NA
    wd = case_when(
      Wind_Direction %in% c("North") ~ 0,
      Wind_Direction %in% c("East") ~ 90,
      Wind_Direction %in% c("South") ~ 180,
      Wind_Direction %in% c("West") ~ 270,
      TRUE ~ NA_real_
    )
  ) %>%
  dplyr::select(ws, wd, Severity)


svg("wind_speed_dir.svg", width = 12, height = 6)

windRose(wind_rose, ws = "ws", wd = "wd", angle = 90,
          paddle = FALSE, statistic = "prop.count", type = "Severity",
         main = expression(bold("Accident Severity by Wind Direction/Speed")),
         key.header   = "Wind Speed (mph)",
         breaks = c(0, 5, 10, 15, Inf),
         key.footer = "",
         cols = alpha(c("#d7d7d7", "#78B7C5", "#E1AF00", "#F21A00"), 0.5),
         annotate = F,
         scales = list(font = 2),
         par.settings=list(fontsize = list(text = 15),
                           par.main.text = list(fontface = "bold"),
                           par.sub.text = list(font = 1)))

dev.off()


```


##### 2: Weather Condition

```{r, fig.height = 5, fig.width = 10}
p2[[7]] <- accidents_trunc %>% 
  drop_na(`Visibility(mi)`) %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  filter(`Visibility(mi)` < 20 & `Visibility(mi)` > 5) %>% 
  ggplot(aes(x = Severity, y = `Visibility(mi)`)) +
  geom_violin(aes(fill = Severity), show.legend = F) +
  labs(x = "", title = "Visibility", y = "Visibility (mi)") +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))


p2[[8]] <- accidents_trunc %>% 
  drop_na(`Humidity(%)`) %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  ggplot(aes(x = Severity, y = `Humidity(%)`)) +
  geom_violin(aes(fill = Severity), show.legend = F) +
  labs(x = "", title =  "Humidity", y = "Humidity (%)") +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))


p2[[9]] <- accidents_trunc %>% 
  drop_na(`Wind_Chill(F)`) %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  ggplot(aes(x = Severity, y = `Wind_Chill(F)`)) +
  geom_violin(aes(fill = Severity), show.legend = F) +
  labs(x = "", title =  "Wind Chill", y = "Wind Chill (F)") +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))


p2[[10]] <- accidents_trunc %>% 
  drop_na(`Temperature(F)`) %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  ggplot(aes(x = Severity, y = `Temperature(F)`)) +
  geom_violin(aes(fill = Severity), show.legend = F) +
  labs(x = "", title =  "Temperature", y = "Temperature (F)") +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))


p2[[11]] <- accidents_trunc %>% 
  drop_na(`Pressure(in)`) %>% 
  mutate(Severity = as.factor(Severity)) %>% 
  ggplot(aes(x = Severity, y = `Pressure(in)`)) +
  geom_violin(aes(fill = Severity), show.legend = F) +
  labs(x = "", title =  "Pressure", y = "Pressure (in)") +
  theme(panel.grid.major.x = element_blank(),
        plot.title = element_text(size = 14)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0")))


(p2[[7]] + p2[[8]]) / 
  (p2[[10]] + p2[[11]]) +
  plot_annotation(title = "How Other Weather Features Impact Accident Severity")


ggsave("weather.svg")
  
```


```{r}

t.test(accidents_trunc %>% filter(Severity == "Severe") %>% drop_na(`Visibility(mi)`) %>% pull(`Visibility(mi)`),
       accidents_trunc %>% filter(Severity == "Minor") %>% drop_na(`Visibility(mi)`) %>% pull(`Visibility(mi)`))

t.test(accidents_trunc %>% filter(Severity == "Severe") %>% drop_na(`Humidity(%)`) %>% pull(`Humidity(%)`),
       accidents_trunc %>% filter(Severity == "Minor") %>% drop_na(`Humidity(%)`) %>% pull(`Humidity(%)`))

t.test(accidents_trunc %>% filter(Severity == "Severe") %>% drop_na(`Temperature(F)`) %>% pull(`Temperature(F)`),
       accidents_trunc %>% filter(Severity == "Minor") %>% drop_na(`Temperature(F)`) %>% pull(`Temperature(F)`))

t.test(accidents_trunc %>% filter(Severity == "Severe") %>% drop_na(`Pressure(in)`) %>% pull(`Pressure(in)`),
       accidents_trunc %>% filter(Severity == "Minor") %>% drop_na(`Pressure(in)`) %>% pull(`Pressure(in)`))


# All significant
```


#### D: Accident Risk by Location


```{r, fig.width = 9, fig.height = 5}
west <- c("WA", "OR", "CA", "NV", "ID", "MT", "WY", "UT", "CO", "AZ", "NM")
south <- c("OK", "TX", "AR", "LA", "MS", "AL", "TN", "KY", "GA", "SC", "FL",
           "NC", "VA", "WV", "DC", "MD", "DE")
midwest <- c("ND", "SD", "NE", "KS", "MN", "IA", "MO", "WI", "IL", "IN",
             "OH", "MI")
northeast <- c("PA", "NY", "NJ", "RI", "CT", "MA", "VT", "NH", "ME")


us_counties <- map_data("county")


county_accidents <- accidents_trunc %>% 
  dplyr::select(Severity, County, State) %>% 
  count(State, County, Severity) %>% 
  mutate(County = tolower(County)) %>% 
  inner_join(census_pop %>% dplyr::select(state, abbr) %>%  distinct(),
             by = c("State" = "abbr")) %>% 
  dplyr::select(state, County, Severity, n) %>% 
  mutate(state = tolower(state)) %>% 
  group_by(state, County) %>% 
  mutate(total = sum(n)) %>% 
  filter(Severity == "Severe") %>%
  mutate(prop = round(n/total * 100, 1)) %>% 
  mutate(prop_bin = case_when(
    prop < 25 ~ "Below 25%",
    prop <= 50 ~ "25% - 50%",
    prop <= 75 ~ "50% - 75%",
    prop > 75 ~ "Above 75%"
  )) %>% 
  mutate(prop_bin = factor(prop_bin, levels = c("Below 25%", "25% - 50%",
                                                "50% - 75%", "Above 75%"))) %>% 
  dplyr::select(state, County, Severity, prop_bin)
  
  
base_county <- us_counties %>%
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", fill = NA) +
  theme(
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  labs(x = "", y = "",
       title = "Share of Severe Crashes by County")  +
  theme(plot.title = element_text(size = 14))
  

counties_crash_map <- us_counties %>%
  inner_join(county_accidents, by = c("region" = "state",
                                      "subregion" = "County"))


p2[[12]] <- base_county +
  geom_polygon(data = counties_crash_map, aes(
  x = long, y = lat, group = group,
  fill = prop_bin
  ), color = "black") +
  scale_fill_manual(values = rev(c("#EA4C46", "#F07470", "#F6BDC0", "#FFEBEA"))) +
  labs(fill = "")


p2[[12]]
  
  
```



```{r, fig.width = 9, fig.height = 5}  

p2[[13]] <- accidents_trunc %>%
  mutate(region = case_when(
  State %in% west ~ "West",
  State %in% south ~ "South",
  State %in% midwest ~ "Midwest",
  State %in% northeast ~ "Northeast"
  )) %>%
  count(region, Severity) %>%
  group_by(region) %>%
  mutate(total = sum(n)) %>%
  ungroup() %>%
  mutate(prop = round(n/total * 100, 1)) %>%
  mutate(region = factor(region, levels = c("Midwest", "Northeast", "South", "West"))) %>%
  ggplot(aes(x = region, y = prop, fill = Severity)) +
  geom_col(position = "stack", color = "black") +
  labs(x = "Region", y = "Proportion", fill = "",
  title = "Proportion of Accident Severity by US Region") +
  scale_y_continuous(labels = label_number(suffix = "%"), expand = c(0, 0.01, 0.11, 0.1)) +
  scale_fill_manual(values = rev(c("#EA4C46", "#F6BDC0"))) +
  theme(plot.title = element_text(size = 14),
  panel.grid.major.x = element_blank())

p2[[13]]

```


```{r, fig.height = 5, fig.width = 15}
p2[[12]] + p2[[13]] +
  plot_annotation("Geographical Analysis of Accident Severity")

ggsave("location.svg")
```


```{r, fig.height = 5, fig.width = 9}
road_infra <- accidents_trunc %>% 
  dplyr::select(Severity, Amenity, Bump, Crossing,
         Give_Way, Junction, No_Exit,
         Railway, Roundabout, Station,
         Stop, Traffic_Calming, Traffic_Signal,
         Turning_Loop)


road_long <- road_infra %>%
  dplyr::select(-Roundabout) %>% 
  rename(`Traffic Signal` = Traffic_Signal) %>% 
  rename(`Traffic Calming` = Traffic_Calming) %>% 
  rename(`No Exit` = No_Exit) %>% 
  rename(`Give Way` = Give_Way) %>% 
  pivot_longer(
    cols      = -Severity,
    names_to  = "feature",
    values_to = "present"
  ) %>%
  filter(present) %>% 
  mutate(feature = factor(feature, levels = rev(c("Junction", "Give Way", "Traffic Calming",
                                              "Railway", "Stop", "Bump", "Amenity",
                                              "No Exit", "Traffic Signal", "Station",
                                              "Crossing", "Turning_Loop"))))

ggplot(road_long, aes(x = feature, fill = Severity)) +
  geom_bar(position = "fill", width = 0.7, color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("Minor" = "#F6BDC0", "Severe" = "#EA4C46")) +
  labs(
    x     = "Road Infrastructure",
    y     = "Proportion of Crashes",
    fill  = "Severity",
    title = "Presence of Road Infrastructure and Crash Severity"
  ) +
  theme(
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(labels = label_number(scale = 100,
    suffix = "%"), expand = c(0, 0, 0.01, 0.01))


ggsave("road_infra.svg")
```


## Part 3: Putting it All Together with a Machine Learning Model

```{r}
ml_data <- accidents_trunc %>% 
  mutate(
    period = hour(Start_Time)) %>% 
  mutate(period = case_when(
    period >= 5 & period <= 11 ~ "Morning",
    period >= 12 & period <= 16 ~ "Afternoon",
    period >= 17 & period <= 20 ~ "Evening",
    period >= 21 | (period <= 4) ~ "Night"
  )) %>% 
  mutate(period = factor(period, 
                         levels = c("Morning", "Afternoon", "Evening", "Night"))) %>% 
  mutate(day_date = wday(Start_Time, abbr = T, label = T)) %>% 
  mutate(day_date = factor(day_date, levels = c("Mon", "Tue", "Wed", "Thu", "Fri",
                                                "Sat", "Sun"))) %>% 
  mutate(month_date = str_extract(month_date, "^[A-Za-z]+")) %>% 
  mutate(month_date = factor(month_date, 
                         levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", 
                                    "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>% 
  mutate(day_date = case_when(
    day_date %in% c("Mon", "Tue", "Wed", "Thu", "Fri") ~ "Weekday",
    day_date %in% c("Sat", "Sun") ~ "Weekend"
  )) %>% 
  mutate(season = case_when(
    month_date %in% c("Dec", "Jan", "Feb") ~ "Winter",
    month_date %in% c("Mar", "Apr", "May") ~ "Spring",
    month_date %in% c("Jun", "Jul", "Aug") ~ "Summer",
    month_date %in% c("Sep", "Oct", "Nov") ~ "Fall"
  )) %>% 
  mutate(season = factor(season, levels = 
                           c("Spring", "Summer", "Fall",
                             "Winter"))) %>%
  mutate(region = case_when(
  State %in% west ~ "West",
  State %in% south ~ "South",
  State %in% midwest ~ "Midwest",
  State %in% northeast ~ "Northeast"
  )) %>% 
  dplyr::select(Severity, period, day_date, season, Wind_Direction, `Wind_Speed(mph)`, Weather_Condition,
         `Temperature(F)`, `Humidity(%)`, `Pressure(in)`, `Visibility(mi)`, region, Amenity, Bump, Crossing,
         Give_Way, Junction, No_Exit, Railway, Station, Stop, Traffic_Calming, Traffic_Signal) %>% 
  drop_na() %>% 
  rename(day = day_date) %>% 
  mutate(Wind_Direction = case_when(
    Wind_Direction %in% c("Calm", "Variable") ~ "Other",
    .default = Wind_Direction
  ))


```


```{r}
set.seed(695)


ml_data <- ml_data %>% 
  group_by(Severity) %>% 
  slice_sample(n = 25000,
               replace = F)


# Train/test split
split <- initial_split(ml_data, prop = c(0.8))

train_data <- split %>%
  training()

test_data <- split %>%
  testing()


sc_train <- scale(train_data[ , c(6, 8, 9, 10, 11)])
train_scaled <- as.data.frame(sc_train)

means <- attr(sc_train, "scaled:center")
sds   <- attr(sc_train, "scaled:scale")

train_data$`Wind_Speed(mph)` <- c(scale(train_data$`Wind_Speed(mph)`))
train_data$`Temperature(F)` <- c(scale(train_data$`Temperature(F)`))
train_data$`Humidity(%)` <- c(scale(train_data$`Humidity(%)`))
train_data$`Pressure(in)` <- c(scale(train_data$`Pressure(in)`))
train_data$`Visibility(mi)` <- c(scale(train_data$`Visibility(mi)`))


test_data$`Wind_Speed(mph)` <- (test_data$`Wind_Speed(mph)` - as.numeric(means[1])) / as.numeric(sds[1])
test_data$`Temperature(F)` <- (test_data$`Temperature(F)` - as.numeric(means[2])) / as.numeric(sds[2])
test_data$`Humidity(%)` <- (test_data$`Humidity(%)` - as.numeric(means[3])) / as.numeric(sds[3])
test_data$`Pressure(in)` <- (test_data$`Pressure(in)` - as.numeric(means[4])) / as.numeric(sds[4])
test_data$`Visibility(mi)` <- (test_data$`Visibility(mi)` - as.numeric(means[5])) / as.numeric(sds[5])

```


```{r}
set.seed(695)

y <- as.numeric(train_data$Severity) - 1
x <- data.matrix(train_data[, -c(1)])

# LASSO - Feature selection
cv_model <- cv.glmnet(x, y, alpha = 1)
best_lambda <- cv_model$lambda.min
plot(cv_model)

# Feature selection: According to best_lambda, we can exclude the '.' variables (these got reduced to 0)
coefficient_model <- coef(glmnet(x, y, alpha = 1, lambda = best_lambda))
selected_features <- rownames(coefficient_model)[coefficient_model@i[-1] + 1]

selected_features
```

```{r}
set.seed(695)

model_logit <- glm(Severity ~ ., data = train_data, family = binomial) %>% 
  stepAIC(direction = "both", trace = F)

summary(model_logit)
vif(model_logit)


prob_predictions <- model_logit %>% 
  predict(test_data, type = "response") # To get probability
predictions <- ifelse(unname(prob_predictions) > 0.5, 1, 0)

observed_predictions <- as.numeric(test_data$Severity) - 1


# Accuracy of 66.2%!
mean(predictions == observed_predictions) 

# AUC Score: 72.1%
auc(observed_predictions, prob_predictions)

# Our model ranks true severe crashes above minor ones 72.1% of the time.

```

```{r}
# Random Forest

set.seed(695)

cl <- makeCluster(detectCores() - 3)
registerDoParallel(cl)


fit_control <- trainControl(
    method = "cv",
    number = 5,
    classProbs = TRUE,
    summaryFunction = twoClassSummary,
    allowParallel = T
  )

train_data$Severity <- factor(train_data$Severity, levels = c("Minor", "Severe"), labels = c("X0", "X1"))
test_data$Severity <- factor(test_data$Severity, levels = c("Minor", "Severe"), labels = c("X0", "X1"))

model_rf <- train(as.factor(Severity) ~ .,
    data = train_data, method = "ranger",
    importance = "permutation",
    trControl = fit_control, verbose = F, tuneLength = 5, metric = "Accuracy" 
  )

stopCluster(cl)


rf_prd_proba <- predict(model_rf, newdata = test_data[, -1], type = ("prob"))
rf_prd <- predict(model_rf, newdata = test_data[, -1])



# 68.5%
mean(test_data$Severity == rf_prd) 


# 75.5%
auc(test_data$Severity, rf_prd_proba[, 2])


```



```{r, fig.width = 9, fig.height = 5}

library(ranger)


train_ranger <- train_data %>% 
  rename(wind_speed = `Wind_Speed(mph)`,
         temp = `Temperature(F)`,
         humidity = `Humidity(%)`,
         pressure = `Pressure(in)`,
         visibility = `Visibility(mi)`)

rf_ranger <- ranger(
  Severity ~ .,
  data       = train_ranger,
  mtry  = 10,
  min.node.size = 1,
  importance = "permutation",  
  probability = FALSE 
)


# 2. pull the vector
imp_tbl <- enframe(
  rf_ranger$variable.importance,
  name  = "feature",
  value = "importance"
) %>%
  arrange(desc(importance))

feature_imp <- imp_tbl %>% 
  head(10) %>% 
  mutate(importance = round(importance / sum(importance) * 100, 2)) %>% 
  mutate(feature = replace(feature, feature == "region", "Region")) %>%
  mutate(feature = replace(feature, feature == "temp", "Temperature")) %>% 
  mutate(feature = replace(feature, feature == "pressure", "Pressure")) %>% 
  mutate(feature = replace(feature, feature == "period", "Period of Day")) %>% 
  mutate(feature = replace(feature, feature == "season", "Season")) %>% 
  mutate(feature = replace(feature, feature == "humidity", "Humidity")) %>% 
  mutate(feature = replace(feature, feature == "Traffic_Signal", "Traffic Signal")) %>% 
  mutate(feature = replace(feature, feature == "day", "Weekday/end")) %>% 
  mutate(feature = replace(feature, feature == "wind_speed", "Wind Speed"))

feature_imp %>%
  ggplot(aes(y = reorder(feature, -importance, desc), x = importance)) +
  geom_col(color = "black", fill = "deepskyblue4", alpha = 0.8) +
  labs(y = "", title = "Feature Importance", x = "Importance (%)") +
  xlim(c(0, 50)) +
  scale_x_continuous(expand = c(0, 0, 0.2, 0.1)) +
  theme(panel.grid.major.y = element_blank())


ggsave("feature_imp.svg")

```


```{r, fig.height = 5, fig.width = 9}
library(ROCR)

pred <- prediction(prob_predictions, observed_predictions)
perf <- performance(pred, "tpr", "fpr")

df_logistic <- data.frame(Model = "Logit (Baseline)", FPR = perf@x.values[[1]], TPR = perf@y.values[[1]])




pred <- prediction(rf_prd_proba[, 2], observed_predictions)
perf <- performance(pred, "tpr", "fpr")

df_rf <- data.frame(Model = "Random Forest", FPR = perf@x.values[[1]], TPR = perf@y.values[[1]])



roc_plot <- ggplot() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  labs(
    x = "False Positive Rate",
    y = "True Positive Rate"
  )




all <- rbind(df_logistic, df_rf) %>%
  mutate(Model = fct(Model,
    levels = c(
      "Logit (Baseline)", "Random Forest")))

logit_plot <- roc_plot +
  geom_line(data = df_logistic, aes(x = FPR, y = TPR), color = "deepskyblue4",
            linewidth = 1.25, alpha = 0.8) +
  ggplot2::labs(
    title = "Logistic Regression (Baseline)",
  ) +
  scale_x_continuous(expand = c(0, 0, 0, 0.01)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(plot.title = element_text(size = 14))


rf_plot <- roc_plot +
  geom_line(data = df_rf, aes(x = FPR, y = TPR), color = "orangered2", 
            linewidth = 1.25, alpha = 0.8) +
  ggplot2::labs(
    title = "Random Forest",
    y = ""
  ) +
  scale_x_continuous(expand = c(0, 0, 0, 0.01)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(plot.title = element_text(size = 14))

roc_plot +
  geom_line(data = all, aes(x = FPR, y = TPR, color = Model, group  = Model),
            linewidth = 1.25, alpha = 0.8) +
  labs(title = "Comparing ROC Curves") +
  scale_color_manual(values = c("deepskyblue4", 
                                "orangered2")) +
  scale_x_continuous(expand = c(0, 0, 0, 0.01)) +
  scale_y_continuous(expand = c(0, 0)) 

# logit_plot + rf_plot +
#   plot_annotation(title = "ROC Curve")


ggsave("roc.svg")


```

